### 2015-10-28 excercise_4/verslag.tex
exercise session 4

\documentclass[11pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{url}
\begin{document}
\section{Interesting knowledge bases}
\begin{enumerate}

\item Musicbrainz
\item \url{data.nobelprize.org}
\item \url{http://datahub.io/dataset/zdb}

\end{enumerate}
\section{Interesting questions}
\begin{enumerate}
\item How many and wich European Parliants members were involved with criminal activities?
\item Wich politicians have good chances of receiving a Nobel Price?
\item Wich song came out on the birthday of a specific politician?
\item Wich politician are murdered.
\end{enumerate}
\end{document}