### 2015-11-02 homework_2/hw2.tex
updated title and cleaned project

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{url}
\usepackage{hyperref}
\author{
  Katrien Laenen\\
  \and
  Gust Verbruggen\\
  \and
  Ward Schodts
}
\title{Knowledge \& the web: homework 2}
\begin{document}
\maketitle
\section*{Which politician have the most tv/radio appearences (on BBC)?}

\end{document}