### 2015-11-02 homework_2/hw2.tex
verslag

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{url}
\usepackage{hyperref}
\usepackage{framed}

\setlength{\parindent}{0pt}

\author{
  Katrien Laenen\\
  \and
  Gust Verbruggen\\
  \and
  Ward Schodts
}
\title{Knowledge \& the web: homework 2}
\begin{document}
\maketitle

The question we chose to answer is

\begin{center}
%\makebox[\textwidth]{%
\bf \Large Which politician has the most television appearances in non-political shows?
%}
\end{center}

\end{document}