### 2016-01-18 project/paper/appendices/googlescholar.tex
added .latex.yml

\section{Queries}
\lstinputlisting[language=SQL, caption=Query 1]{../queries/allnobelwinners.rq}