### 2016-01-23 project/paper/appendices/sparql.tex
added appendix C

\section{Linked open data}
\label{sec:lod}
\subsection{Sparql endpoints}
\begin{table}[H]
\centering
\begin{tabular}{l|l}
	\textbf{\textsc{Source}} & \textbf{\textsc{Sparql endpoint}} \\ \hline
	\rule{0pt}{4mm}DBPedia &  \\
	Year of birth & \textbf{DBPedia} \\
	University ranking & \textbf{DBPedia} + Shanghai Ranking 2015\\
	Work Productivity & Google Scholar, \textbf{ToE}\\
	Popularity & Facebook\\
	Nobel Prize& \textbf{Nobelprize.org}
\end{tabular}
\caption{Overview of used features and their respective sources.}
\end{table}
\subsection{Difficulites with LoD}
In this section we discuss our difficulties that we experienced with linked open data (LoD). These are worth mentioning because they influenced our results.

\subsubsection{The language barrier}
The great thing about LoD is that it comes in different languages, what means that you have a bigger and a broader information source.

\subsubsection{LoD is sometimes broken}