### 2016-01-23 project/paper/appendices/googlescholar.tex

\section{Retrieving data from Google Scholar}
\label{sec:googlescholar}

Retrieving data from Google Scholar is a laborous task.
This appendix provides our experience and also our tips on how to scrape Google Scholar.