### 2016-01-26 project/paper/sections/methodology.tex
extra stap in methodology

\section{Methodology}
\label{sec:methodology}
To answer the research question we will use the approach described below. The next sections will explain the steps more thoroughly.
\begin{enumerate}
\item We decide which persons we exactly want in the training dataset and which persons we want in our research dataset.
\item We think about what properties of previous Nobel Prize winners might be interesting to incorporate in learning a model. This is the feature selection step. Furthermore, we determine whether and where this information can be found.
\item The available data is then collected by the aid of SPARQL queries and scraping webpages. This is combined into a training dataset and a research dataset. The training dataset can be used to learn a predictive model. The research dataset is used to answer the research question.
\item Some statistics are applied to the training dataset to assess its quality. For example, because collecting the data is prone to errors, some outlier detection and removal has to be performed.
\item The final training dataset is used to learn a model that is able to predict how likely a person is to win a Nobel Prize. A logistic regression model is used. Because it predicts the probability of particular outcomes, it is perfectly suited to answer the research question. The model's accuracy is assessed using cross-validation on the training dataset.
\item The final step is applying the learned model on the research dataset and discussing the results. 
\end{enumerate}